---
title: "An analysis of the spotify dataset"
author: "Sharmini Joseph Pereira (S2114594) | Kristian Surya Dinata (S2043845)"
date: "5/8/2022"
output: word_document
---

#1 Data Preparation

## The data for this analysis comes from kaggle, source: https://www.kaggle.com/datasets/pepepython/spotify-huge-database-daily-charts-over-3-years?resource=download

## Library used to clean the data
```{r}
library(dplyr)
library(tidyr)
library(janitor)
library(tidyverse)
library(tibble)
library(grid)
library(ggplot2)
library(knitr)
library(kableExtra)
```

## Importing the dataset which is obtained from Github 
```{r}
url = "https://raw.githubusercontent.com/Kristian2828/pfds_group_project/main/dataset/dataset.csv"
df1 = read.csv(url)
```

## Description of the dataset
```{r}
#dimension of the dataset
yes_0 = str_trim(paste0(dim(df1), sep = "  ", collapse = ""))

#column names of the dataset
vars_0 = str_trim(paste0(names(df1), sep = " | ", collapse = ""))

# Creating a table
table_attributes = data_frame(Data = 'Spotify_Dataset_BeforeCleaning',
  `Rows  Columns` = yes_0,
  Variables = vars_0)

# Printing the table
kable(table_attributes, format = "html") %>%
  kable_styling(bootstrap_options = "striped") %>%
     column_spec(2, width = "12em")
```
## Exploration of dataset
```{r}
# Check the dimensions of the data set
dim = dim(df1)

# View the structure of the data
glimpse(df1)
 
# View the column names because these will be useful handles
names(df1)

# View a summary of the data
summary(df1)

# View first 6 rows
head(df1)

# View last 6 rows
tail(df1)
```

#2. Cleaning the data
```{r}
# Format variable names
df2 = clean_names(df1)
View(df2)

# Removing incomplete observations
# Removing NA values
df2 = na.omit(df2)

# Removing irrelevant columns
df2 = df2 %>% select(-c(uri,key,genre_new,released_after_2017,explicit_false,explicit_true,album_2))
df3 = df2[-c(26:66,88:140)]
```

## box plot of the audio features to find any extreme outliers
```{r}
boxplot(df3$loudness,
  ylab = "loudness",
  main = "Boxplot of loudness",
  ylim = c(-30, 0)
)

boxplot(df3$danceability,
  ylab = "dancebility",
  main = "Boxplot of dancebility"
)
 
boxplot(df3$energy,
  ylab = "energy",
  main = "Boxplot of energy"
)

boxplot(df3$acoustics,
  ylab = "acoustics",
  main = "Boxplot of acoustics"
)

boxplot(df3$tempo,
 ylab = "tempo",
  main = "Boxplot of tempo"
)

#It's concluded there is no extreme outliers found in audio features
```
#3. Data manipulation
#duration_minutes - which will have the duration of the song in minutes
```{r}
#Creating a duration in minutes column
df3$duration_minutes <- df3$duration_ms/(1000*60)
```
#4. Finishing touches
```{r}
#dimension of the dataset
yes = str_trim(paste0(dim(df3), sep = "  ", collapse = ""))

#column names of the dataset
vars = str_trim(paste0(names(df3), sep = " | ", collapse = ""))

# Creating a table
table_attributes = data_frame(Data = 'Spotify_Dataset_AfterCleaning',
  `Rows  Columns` = yes,
  Variables = vars)

# Printing the table
kable(table_attributes, format = "html") %>%
  kable_styling(bootstrap_options = "striped") %>%
     column_spec(2, width = "12em")
```
## Exporting the dataset
```{r}
write.csv(df3, "data_cleaned.csv")
```

